\section{Experiments}
\label{sec:experiments}

\paragraph{Benchmarks.}
We evaluate on the three-benchmark suite standard for LLM agent SFT
\citep{song2024eto,zhang2024ipr,zhang2025alee}.
\textbf{ALFWorld} \citep{shridhar2021alfworld} is a text-based household
environment (six task categories; binary success).
\textbf{ScienceWorld} \citep{wang2022scienceworld} is a science simulator with
30 task types requiring causal reasoning over 10--120 steps
(average normalised score $\times100$); the unseen split is the primary
generalisation target.
\textbf{WebShop} \citep{yao2022webshop} requires sequential product search over
12,087 items (average reward $\times100$).

\paragraph{Data.}
All experiments use the \texttt{agent-eto/eto-sft-trajectory} dataset
\citep{song2024eto}: expert ReAct trajectories with GPT-4 rationales.
No environment interaction or reward labelling is performed for any variant;
all methods train on identical data.

\paragraph{Baselines.}
\textbf{Zero-shot LLaDA~2.0-mini}: untuned base model.
\textbf{Standard SFT}: a sample-efficient baseline that mimics autoregressive
next-action prediction. Each trajectory is sliced into multiple examples,
where each action $A_t$ is masked and predicted from its full causal history
$\{O_0, A_0, \ldots, O_t\}$.
\textbf{Prefix SFT (offline IWM)}: ALEE's IWM objective adapted to an offline,
two-stage prefix-conditioning task.
Stage~1 (IWM) predicts $O_{t+1}$ from the history and $(O_t, A_t)$.
Stage~2 (Policy SFT) is then fine-tuned to predict $A_t$ from the history and $O_t$.
\textbf{\aomt{}-Action-Only}: mask restricted to action units, with the full
trajectory as bidirectional context.
\textbf{\aomt{}-Mixed} (full method): each action and observation unit
(excluding the initial objective) is masked independently with probability $p$.
ETO \citep{song2024eto}, IPR \citep{zhang2024ipr}, and Early Experience IWM
\citep{zhang2025alee} are cross-paradigm references requiring rollouts
(and rewards for ETO/IPR).

\paragraph{Model and training.}
All experiments use LLaDA~2.0-mini \citep{bie2025llada20}: 16B total / 1.4B
active parameters (MoE). Trajectories are tokenised as flat sequences with
separator tokens; unit spans are recorded at preprocessing so the mask sampler
enforces whole-unit masking. Hyperparameters: lr $2.5\times10^{-5}$ (cosine
decay to $2.5\times10^{-6}$), 50 warmup steps, weight decay 0.1, gradient
clip~1.0, batch size~64, sequence length~2048, bf16. For a fair comparison, all
experiments are trained for the same total number of gradient updates,
normalised against 3 epochs over the original unprocessed dataset.
Mask probability $p$ is swept
over $\{0.15, 0.25, 0.40, 0.50\}$ on ALFWorld and transferred to other
benchmarks. Inference: iterative unmasking, temperature~0.0, block length~32,
32 decoding steps \citep{bie2025llada20}.

\paragraph{Evaluation metrics.}
Primary metrics are task success~(\%) for ALFWorld, average normalised score for
ScienceWorld, and average reward for WebShop. We additionally report
\emph{observation-masked NLL}:
\begin{equation}
  \NLL_{\text{obs}}(\theta) =
    -\mathbb{E}_{\tau \sim \Dcal_{\text{test}}}\;
    \mathbb{E}_{t}\!\left[\log p_\theta(O_t \mid \tau \setminus O_t)\right],
  \label{eq:nll_obs}
\end{equation}
where $\tau \setminus O_t$ provides all other units as clean bidirectional
context. $\NLL_{\text{obs}}$ is a pseudo-log-likelihood
\citep{salazar2020masked} comparable across masked DLMs but incommensurable
with autoregressive log-likelihoods (Appendix~\ref{app:nll_details}).
Standard SFT and Prefix SFT are excluded from this metric.
Robustness is evaluated by randomly corrupting fraction
$\rho \in \{0.1, 0.2, 0.3\}$ of observation tokens at test time
(ALFWorld seen).
