\section{Conclusion}
\label{sec:conclusion}

We introduced Any-Order Masked Training (\aomt{}), a reward-free offline SFT
paradigm for LLM-based agents that treats training-order assignment as a
variable rather than a fixed commitment. By resampling unit-level masks across
epochs and reconstructing masked trajectory spans from bidirectional context,
\aomt{} extracts a richer family of supervisory signals from a fixed expert
corpus than any single-order method. We proved that standard SFT and offline
ALEE-IWM are degenerate special cases under constrained mask samplers, and
introduced observation-masked NLL as a natural evaluation metric for implicit
world-model quality.
\PLACEHOLDER{Summarise key empirical findings: (1) \aomt{}-Mixed outperforms SFT
across all benchmarks; (2) bidirectional context and world-model training
contribute independently; (3) $\NLL_{\mathrm{obs}}$ correlates with task success
on ScienceWorld but not WebShop; (4) \aomt{}-Mixed degrades more gracefully under
observation noise.}

\paragraph{Limitations and future work.}
\aomt{} currently operates on text-based trajectories; extension to visual
observations requires unit-level masking over image patches, building on
multimodal masked DLMs \citep{zhou2025lladav}. The formal relationship between
\aomt{}'s unit-level objective and LLaDA's ELBO warrants further analysis
(Appendix~\ref{app:elbo}). The most immediate extension is composing \aomt{}
with ALEE's data-collection pipeline---applying any-order masked reconstruction
to agent-generated rollouts to address both data-diversity and objective-richness
bottlenecks simultaneously.

\paragraph{Broader impact.}
\aomt{} improves sample efficiency of SFT on fixed offline corpora without
requiring additional environment interaction or reward labelling. Risks are those
generic to improved LLM-agent capabilities.
