\section*{Appendix}
\setcounter{section}{0}
\renewcommand{\thesection}{\Alph{section}}

\section{Proofs}
\label{app:proofs}

\begin{proof}[Proof of Proposition~\ref{prop:subsumption}]
\textbf{(i) Standard SFT.}
Fix $m$ to always mask exactly one action unit $A_t$, with context
$\Ccal(m) = \{O_0, A_0, \ldots, O_t\}$ and a causal attention mask preventing
any future unit from contributing to the reconstruction of $A_t$.
Substituting into Eq.~\eqref{eq:aomt}:
\[
  \Lcal_{\text{AOMT}}(\tau,m) = -\log p_\theta(A_t \mid O_0, A_0, \ldots, O_t),
\]
which is exactly the standard SFT loss summed over $t = 0,\ldots,T-1$.

\textbf{(ii) Offline ALEE-IWM.}
Fix $m$ to always mask $O_{t+1}$, with context $\Ccal(m) = \{O_t, A_t\}$
(local context only).
Substituting into Eq.~\eqref{eq:aomt}:
\[
  \Lcal_{\text{AOMT}}(\tau,m) = -\log p_\theta(O_{t+1} \mid O_t, A_t),
\]
which is the ALEE-IWM loss restricted to the offline expert corpus.
\aomt{} with an unconstrained Bernoulli($p$) sampler strictly generalises both
by allowing any subset of units to be masked and using full bidirectional
attention over all context units.
\end{proof}

\section{Relationship to Masked Diffusion ELBO}
\label{app:elbo}

LLaDA \citep{nie2025llada} trains with:
\begin{equation}
  \Lcal_{\text{LLaDA}} = -\mathbb{E}_{t \sim U[0,1]}\,
    \mathbb{E}_{m \sim \mathrm{Mask}(t)}\,
    \frac{1}{t} \sum_{i : m_i=1} \log p_\theta(x_i \mid x \odot (1-m)),
\end{equation}
a variational upper bound on token-level NLL.
\aomt{} replaces token-level masking with unit-level masking and fixes the
schedule to a single probability $p$, trading the ELBO interpretation for a
cleaner semantic objective aligned with agent decision granularity.
We conjecture that \aomt{} inherits Fisher consistency from LLaDA at the unit
level; formal proof is left to future work.

At inference, \aomt{} uses LLaDA~2.0's block diffusion decoding (iterative
unmasking over 32 steps), not single-pass reconstruction. The any-order
learning effect emerges from mask diversity \emph{across training steps},
not within-trajectory iterative refinement.

\section{Observation-Masked NLL: Implementation Details}
\label{app:nll_details}

As LLaDA predicts each masked token independently in a single forward pass,
$\NLL_{\text{obs}}$ (Eq.~\eqref{eq:nll_obs}) is a \emph{pseudo-log-likelihood}
in the sense of \citet{salazar2020masked}: it sums per-token log-probabilities
assuming conditional independence of tokens in $O_t$ given context, rather than
computing the true joint $\log p(O_t \mid \tau \setminus O_t)$. It is directly
comparable across masked DLM checkpoints but not commensurable with
autoregressive log-likelihoods, which condition only on past context and
therefore measure forward prediction rather than bidirectional world-model quality.

Standard SFT and Prefix SFT Stage~1 are excluded from $\NLL_{\text{obs}}$
tables: the former has no mechanism for future conditioning; the latter was
trained with single-step context $(O_t, A_t)$ only, making full-trajectory
evaluation out-of-distribution.

\section{Attention Heatmaps}
\label{app:attention}

\PLACEHOLDER{Final-layer attention weights averaged across all heads over 50
held-out ALFWorld trajectories, when reconstructing a masked mid-trajectory
action. Left: \aomt{}-Mixed. Right: Standard SFT (zero-future-attention
reference). AOMT-Mixed is expected to show substantial weight on future context
units; Standard SFT should show none. Generated after experiments complete.}

\section{Failure Mode Analysis}
\label{app:failures}

\PLACEHOLDER{Table of 3--5 representative cases where \aomt{}-Mixed succeeds
and Standard SFT fails on ScienceWorld unseen tasks. For each case: task
description, step at which SFT diverges, SFT's incorrect action,
\aomt{}-Mixed's correct action, and the key piece of bidirectional context
that distinguishes them.}

\section{Full Results by Task Category}
\label{app:full_results}

\PLACEHOLDER{Per-category breakdown of ALFWorld results (all six task types)
and ScienceWorld results (all 30 task types) for all offline methods.}
